Data Analysis

```{r setup}
# Load your packages
library(tidyverse); library(here); library(lubridate); library(sf); library(mapview); library(leaflet); library(trend); library(Kendall)

# Set your working directory
here()

#Set theme
my_theme <-  theme(
    plot.title = element_text(face = "bold", size = 12),
    panel.background = element_rect(fill = "white", colour = NA), 
    panel.border = element_rect(fill = NA, colour="grey50"), 
    panel.grid.major = element_line(colour = "black", size = 0.01),
    panel.grid.minor =  element_line(colour = "black", size = 0.01),
    axis.text = element_text(size = 10),
    axis.ticks = element_blank(),
  )

#Load data
obis_nc <- read.csv(here('Data/Processed/obis_NC_2000-2019.csv'))
```

Temporal Analysis: Annual Trends
```{r year graph}
obis_nc$year <- as.factor(obis_nc$year)
obis_nc$month <- as.factor(obis_nc$month)

obis_nc %>%
  group_by(year) %>%
  summarise(individuals = sum(group_size)) %>%
  ggplot(aes(x=year, y=individuals)) +
  geom_bar(stat="identity") +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  labs(title = "Annual Sightings of Cownose Rays in North Carolina (2000-2019)",
       x = "Year", y = "Number of Individuals") +
  my_theme
```

```{r}
obis_year <- obis_nc %>%
  group_by(year) %>%
  summarise(individuals = sum(group_size))

obis_year$year <- as.numeric(as.character(obis_year$year))

obis_year %>%
  ggplot(aes(x = year, y = individuals)) +
  geom_line() +
  geom_smooth(method=lm, se=FALSE) +
  labs(title = "Annual Sightings of Cownose Rays in North Carolina (2000-2019)",
       x = "Year", y = "Number of Individuals") +
  my_theme
```
```{r}
year.regression <- lm(data = obis_year, individuals ~ year)
summary(year.regression)
```
Significant increase in sightings between 2000 and 2019 as p-value <0.05, but R-squared suggests only 29.2% of variability is explained by year variable.


Temporal analysis: monthly trends
```{r month graph}
obis_month <- obis_nc %>%
  group_by(month) %>%
  summarise(individuals = sum(group_size)) 

month <- 1:2
individuals <- 0
df <- data.frame(month,individuals)
obis_month_complete <- rbind(df, obis_month)

obis_month_complete %>%
  ggplot(aes(x=month, y=individuals)) +
  geom_bar(stat="identity") +
  labs(title = "Monthly Sightings of Cownose Rays in North Carolina (2000-2019)",
       x = "Month", y = "Number of Individuals") +
  scale_x_discrete(limits=c('1','2','3','4','5','6','7','8','9','10','11','12'),
                   breaks = seq_along(month.abb), 
                   labels = month.abb) +
  my_theme
```

```{r}
obis_month_complete$month <- as.numeric(as.character(obis_month_complete$month))

obis_month_complete %>%
  ggplot(aes(x = month, y = individuals)) +
  geom_line() +
  geom_smooth(method=lm) +
  scale_x_continuous(n.breaks = 12) +
  labs(title = "Monthly Sightings of Cownose Rays in North Carolina (2000-2019)",
       x = "Month", y = "Number of Individuals") +
  my_theme
```
```{r}
month.regression <- lm(data = obis_month_complete, individuals ~ month)
summary(month.regression)
```

GLM cannot determine monthly trends, continue using time series analysis

Time Series Analysis
```{r time series clean}
obis_summary <- obis_nc %>%
  group_by(year,month) %>%
  summarize(individuals = sum(group_size))
obis_summary$date = ymd(
  paste(obis_summary$year, obis_summary$month, "1"))

days <- data.frame(seq(as.Date("2000-01-01"),as.Date("2019-12-01"),by="1 month"))
colnames(days) <- "date"

obis_summary <- left_join(days, obis_summary)
obis_summary <- obis_summary %>%
  mutate(year = year(date), month = month(date)) %>%
  mutate_all(~replace(., is.na(.), 0))
```

```{r time series analysis}
f_month <- first(obis_summary$month)
f_year <- first(obis_summary$year)

obis_ts <- ts(obis_summary$individuals,
                   start=c(f_year,f_month),
                   frequency=12) 

decomp <- stl(obis_ts,s.window = "periodic")
plot(decomp)
```

```{r kendall}
monthly_trend <- Kendall::SeasonalMannKendall(obis_ts)
summary(monthly_trend)
```

```{r}
monthly_trend2 <- trend::smk.test(obis_ts)
summary(monthly_trend2)
```
Although the overall seasonal Mann-Kendall test had a p-value >0.05, indicating there is no overall monthly trend, breaking the time series analysis down into individual months revealed that the month of June has a significantly different number of sighings than other months as p-value was 0.016. The month of May had a p-value of 0.056, which is very close to the threshold of p<0.05, and suggests a possible significant difference for that month as well that warrants further investigation.


Spatial Analysis
```{r}
mapview(nc_counties,
        col.regions = 'lightyellow', 
        alpha.regions = 0.2, 
        map.types = "CartoDB.Positron") +
  mapview(obis_sf, 
          cex = "group_size", legend=TRUE,
          zcol = "month", col.regions=list('red','blue')
          )
```

Extract specific years with most sightings to explore annual movement patterns
```{r}
obis_2015 <- obis_sf %>%
  filter(year==2015)
obis_2016 <- obis_sf %>%
  filter(year==2016)
obis_2018 <- obis_sf %>%
  filter(year==2018)
obis_2019 <- obis_sf %>%
  filter(year==2019)
```

Toggle year layers on and off to see specific year
```{r}
mapview(nc_counties,
        col.regions = 'lightyellow', 
        alpha.regions = 0.2, 
        map.types = "CartoDB.Positron") +
  mapview(obis_2015, 
          cex = "group_size", legend=TRUE) +
  mapview(obis_2016, 
          cex = "group_size", legend=TRUE) +
  mapview(obis_2018, 
          cex = "group_size", legend=TRUE) +
  mapview(obis_2019, 
          cex = "group_size", legend=TRUE)
```

Extract by month to see monthly movement patterns
```{r}
Mar <- obis_sf %>%
  filter(month==3)
Apr <- obis_sf %>%
  filter(month==4)
May <- obis_sf %>%
  filter(month==5)
Jun <- obis_sf %>%
  filter(month==6)
Jul <- obis_sf %>%
  filter(month==7)
Aug <- obis_sf %>%
  filter(month==8)
Sep <- obis_sf %>%
  filter(month==9)
Oct <- obis_sf %>%
  filter(month==10)
Nov <- obis_sf %>%
  filter(month==11)
Dec <- obis_sf %>%
  filter(month==12)
```

toggle month layers on and off to see specific month
```{r}
mapview(nc_counties,
        col.regions = 'lightyellow', 
        alpha.regions = 0.2, 
        map.types = "CartoDB.Positron") +
  mapview(Mar, cex = "group_size", legend=TRUE) +
  mapview(Apr, cex = "group_size", legend=TRUE) +
  mapview(May, cex = "group_size", legend=TRUE) +
  mapview(Jun, cex = "group_size", legend=TRUE) +
  mapview(Jul, cex = "group_size", legend=TRUE) +
  mapview(Aug, cex = "group_size", legend=TRUE) +
  mapview(Sep, cex = "group_size", legend=TRUE) +
  mapview(Oct, cex = "group_size", legend=TRUE) +
  mapview(Nov, cex = "group_size", legend=TRUE) +
  mapview(Dec, cex = "group_size", legend=TRUE)
```